I/O simulation model with Linux page cache, integration and evaluation in SimGrid framework

### 2020-10-23 paper_io_sim.tex
Update figures

			In simulated platform description, because the disk bandwidths in SimGrid 
			are symmetrical, we measure the read/write bandwidths on our cluster and 
			take the average bandwidths. 
			As the pipelines are single-core, we create simulated tasks, 
			each of them runs on only one core. 
			While in Python, computation time can be directly configured, 
			this time is injected to the WRENCH simulators with number of flops, 
			which is calculated with CPU speed and compute duration. 
			In our experiments, the CPU speed is set to 1 Gflops. 
			Finally, because WRENCH simulates applications base on network-communication, 
			we use an infinite bandwidth to eliminate network latency for local I/Os.
			The details of simulated platform is available here: 
			https://github.com/dohoangdzung/wrench/blob/io_chunk_test/examples/basic-examples/io-pagecache/single_host.xml

@@

			Memory behaviors of real pipeline execution and simulators are recorded 
			and the memory profiling results with 20GB and 100GB input files 
			are illustrated in Figure~\ref{fig:single_memprof_20gb} and Figure~\ref{fig:single_memprof_100gb}. 
			The bars in the figures present activities in tasks (read, compute and write), 
			while the lines describe memory status throughout run time of the pipeline. 
			Because the results with 20Gb, 50GB, 100GB are analogous, we only include 
			the results with 20GB and 100GB in this paper.

@@

			The I/O time of tasks in simulators and real execution is also tracked and compared. 
			Figure~\ref{fig:single_error_20} and Figure~\ref{fig:single_error_100} show 
			the relative error of simulated I/O time of individual I/O activities. 
			Our pipeline consists of 3 tasks, and each task reads an input file, 
			performs some computation, and writes output file. 
			Thus, there are 6 sequential I/O activities, \textit{read1} and \textit{write1} 
			from the first task, \textit{read2} and \textit{write2} the second task, 
			and \textit{read3}, \textit{write3} from the last task.

@@

			\begin{figure*}
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_20gb.pdf}
   				\caption{Memory profiling with 20GB input file}	\label{fig:single_memprof_20gb}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_100gb.pdf}
   				\caption{Memory profiling with 100GB input file}	\label{fig:single_memprof_100gb}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/error_20.pdf}
   				\caption{Simulation error with 20GB input file}	\label{fig:single_error_20}
			\end{subfigure}		
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/error_100.pdf}
   				\caption{Simulation error with 100GB input file}	\label{fig:single_error_100}
			\end{subfigure}	
			\caption{Simulation results with different input file sizes}
			\label{fig:single_results}
			\end{figure*}

			\begin{figure*}
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/fincore_real.pdf}
   				\caption{Real pipeline execution}	
   				\label{fig:fincore_real}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/fincore_wrench.pdf}
   				\caption{WRENCH with page cache simulator}	
   				\label{fig:fincore_wrench}
			\end{subfigure}	
			\caption{Amount of file data in cache after each I/O activity with 100GB of input}
			\label{fig:fincore}
			\end{figure*}

			This longer read can be explained with the amount of cached data of files 
			at the beginning of the read in reality and simulators shown in Figure~\ref{fig:fincore}. 			
			In this figure, \textit{file$<$i$>$} is the input of \textit{read$<$i$>$} and 
			output of \textit{write$<$i-1$>$}. 
			In the real execution, after \textit{write2}, which writes \textit{file3}, 
			all data of \textit{file3} is in cache (Figure~\ref{fig:fincore_real}). 
			Thus, in \textit{read3}, \textit{file3} is read with memory bandwidth. 
			But in our simulator, after \textit{write2}, only a part of \textit{file3} 
			remains in cache (Figure~\ref{fig:fincore_wrench}). 
			The reason is while \textit{file3}  is being written, the page cache is saturated, 
			eviction is then evoked to evict data from page cache, make space 
			available for \textit{file3}. 
			However, because the evicted amount is less than required, 
			some \textit{file3} data is evicted after being written, make \textit{file3} 
			partially cached. 
			As a result, in \textit{read3}, only a part of \textit{file3} is read from cache 
			with memory bandwidth, a part of it is read from disk with disk read bandwidth, 
			which is many times slower, leading to longer read time.

@@

			\begin{figure*}
			\begin{subfigure}{0.33\linewidth}
				\centering
   				\includegraphics[width=\linewidth]{result/multi/figures/makespan.pdf}
   				\caption{Total makespan}	
   				\label{fig:multi_makespan}
			\end{subfigure}	
			\begin{subfigure}{0.33\linewidth}
				\centering
   				\includegraphics[width=\linewidth]{result/multi/figures/readtime.pdf}
   				\caption{Average read time}	
   				\label{fig:multi_readtime}
			\end{subfigure}	
			\begin{subfigure}{0.33\linewidth}
				\centering
   				\includegraphics[width=\linewidth]{result/multi/figures/writetime.pdf}
   				\caption{Average write time}	
   				\label{fig:multi_writetime}
			\end{subfigure}		
			\caption{Total makespan and I/O time with different number of concurrent pipeline}
			\label{fig:multi_result}
			\end{figure*}

			In this experiment, the results is analyzed in terms of the total makespan  
			and average I/O time concurrent pipelines.
			The results from real execution and simulation are presented in Figure~\ref{fig:multi_result}.

			The Figure~\ref{fig:multi_makespan} shows that as the number of pipelines 
			increases, the total makespan raises accordingly because the amount of data 
			increases but the memory and disk bandwidths are fixed and shared between pipelines. 
			In comparison with original WRENCH, the result from WRENCH with page cache are closer 
			to reality. 
			This proximity can also be seen in average read time shown in 
			Figure~\ref{fig:multi_readtime}. 
			Especially, when it comes to the average write time 
			(shown in Figure~\ref{fig:multi_writetime}), 
			although the trend in reality is more complex, our WRENCH with page cache 
			simulator can still capture it and share the same pattern with 
			the real execution, when they gradually increase before surging 
			with the same slope after the number of pipelines reaches 10. 
			This can be explained that with less than 10 concurrent pipelines, 
			the page cache is not saturated, so all files can be written entirely to 
			cache with memory bandwidth in a short time. 
			After the page cache is saturated with dirty data (at around 
			10 concurrent pipelines in this experiment), this dirty data needs 
			to be flushed in order to make space available for writing new data to cache. 
			This dirty data is flushed to disk with disk bandwidth, which is much  
			slower than memory bandwidth. 
			The more pipelines we have, the more data needs to be written, the more data 
			needs to be flushed, leading to a sharper increase in the average write time.

>>>>

			In simulated platform description, because the disk bandwidths in SimGrid 
			are symmetrical, we measure the read/write bandwidths on our cluster and 
			take the average bandwidths. 
			As the pipelines are single-core, we create simulated tasks, 
			each of them runs on only one core. 
			While in Python, computation time can be directly configured, 
			this time is injected to the WRENCH simulators with number of flops, 
			which is calculated with CPU speed and compute duration. 
			In our experiments, the CPU speed is set to 1 Gflops. 
			Finally, because WRENCH simulates applications base on network-communication, 
			we use an infinite bandwidth to eliminate network latency for local I/Os.
			The details of simulated platform is available here: 
			\url{https://github.com/dohoangdzung/wrench/blob/io_chunk_test/examples/basic-examples/io-pagecache/single_host.xml}

@@

			Memory behaviors of real pipeline execution and simulators are recorded 
			and the memory profiling results with 20GB and 100GB input files 
			are illustrated in Figure~\ref{fig:single_memprof}. 
			The bars in the figures present activities in tasks (read, compute and write), 
			while the lines describe memory status throughout run time of the pipeline. 
			Because the results with 20Gb, 50GB, 100GB are analogous, we only include 
			the results with 20GB and 100GB in this paper.

@@

			The I/O time of tasks in simulators and real execution is also tracked and compared. 
			Figure~\ref{fig:single_errors} shows the relative error of simulated I/O time 
			of individual I/O activities. 
			Our pipeline consists of 3 tasks, and each task reads an input file, 
			performs some computation, and writes output file. 
			Thus, there are 6 sequential I/O activities, \textit{read1} and \textit{write1} 
			from the first task, \textit{read2} and \textit{write2} the second task, 
			and \textit{read3}, \textit{write3} from the last task.

@@

			\begin{figure*}
			\centering
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_20gb.pdf}
   				\caption{20GB input file}	\label{fig:single_memprof_20gb}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_100gb.pdf}
   				\caption{100GB input file}	\label{fig:single_memprof_100gb}
			\end{subfigure}	
			\caption{Memory profiling results with different input file sizes}
			\label{fig:single_memprof}	
			\begin{subfigure}{\linewidth}
				\centering
   				\includegraphics[width=\linewidth]{result/single/figures/errors.pdf}
			\end{subfigure}
			\caption{Simulation errors with different input file sizes}
			\label{fig:single_error}	
			\begin{subfigure}{\linewidth}
				\centering
   				\includegraphics[width=\linewidth]{result/single/figures/cached_files.pdf}
			\end{subfigure}	
			\label{fig:single_cache}
   			\caption{Amount of file data in cache after each I/O activity with 100GB of input}
			\end{figure*}

			This longer read can be explained with the amount of cached data of files 
			at the beginning of the read in reality and simulators shown in 
			Figure~\ref{fig:single_cache}. 			
			In this figure, \textit{file$<$i$>$} is the input of \textit{read$<$i$>$} and 
			output of \textit{write$<$i-1$>$}. 
			In the real execution, after \textit{write2}, which writes \textit{file3}, 
			all data of \textit{file3} is in cache (Figure~\ref{fig:fincore_real}). 
			Thus, in \textit{read3}, \textit{file3} is read with memory bandwidth. 
			But in our simulator, after \textit{write2}, only a part of \textit{file3} 
			remains in cache (Figure~\ref{fig:fincore_wrench}). 
			The reason is while \textit{file3}  is being written, the page cache is saturated, 
			eviction is then evoked to evict data from page cache, make space 
			available for \textit{file3}. 
			However, because the evicted amount is less than required, 
			some \textit{file3} data is evicted after being written, make \textit{file3} 
			partially cached. 
			As a result, in \textit{read3}, only a part of \textit{file3} is read from cache 
			with memory bandwidth, a part of it is read from disk with disk read bandwidth, 
			which is many times slower, leading to longer read time.

@@

			\begin{figure*}		
			\begin{subfigure}{\linewidth}
				\centering
   				\includegraphics[width=\linewidth]{result/multi/figures/multi_local.pdf}
			\end{subfigure}		
			\caption{I/O time with different numbers of concurrent pipeline}
			\label{fig:multi_local}
			\end{figure*}

			In this experiment, the results is analyzed with the average I/O time of concurrent pipelines.
			The results from real execution and simulation are presented in Figure~\ref{fig:multi_result}.

			The Figure~\ref{fig:multi_readtime} shows that as the number of pipelines 
			increases, the average read time raises accordingly because the amount of data 
			increases but memory and disk bandwidths are fixed and shared between pipelines. 
			In comparison with original WRENCH, the result from WRENCH with page cache are closer 
			to reality. 
			Especially, when it comes to the average write time 
			(shown in Figure~\ref{fig:multi_writetime}), 
			although the trend in reality is more complex, our WRENCH with page cache 
			simulator can still capture it and share the same pattern with 
			the real execution, when they gradually increase before surging 
			with the same slope after the number of pipelines reaches 10. 
			This can be explained that with less than 10 concurrent pipelines, 
			the page cache is not saturated, so all files can be written entirely to 
			cache with memory bandwidth in a short time. 
			After the page cache is saturated with dirty data (at around 
			10 concurrent pipelines in this experiment), this dirty data needs 
			to be flushed in order to make space available for writing new data to cache. 
			This dirty data is flushed to disk with disk bandwidth, which is much  
			slower than memory bandwidth. 
			The more pipelines we have, the more data needs to be written, the more data 
			needs to be flushed, leading to a sharper increase in the average write time.
